\documentclass[a4paper]{article}
\usepackage{stdtemplate}

\title{MAT347 Abstract Algebra}
\author{Jonah Chen}
\date{}
\begin{document}
\maketitle
\sffamily
\section{Groups}
Groups are generally associated with symmetries. Consider the equilateral triangle:
\begin{center}
    \begin{tikzpicture}
        % draw a triangle and label the vertices A, B, C
        \draw (0,0) -- (4,0) -- (2,3.464) -- cycle;
        \node[above] at (2,3.464) {$A$};
        \node[below] at (4,0) {$B$};
        \node[below] at (0,0) {$C$};
    \end{tikzpicture}
\end{center}
We know that there are six symmetries of the triangle:
\begin{itemize}
    \item Identity transformation (do nothing) denoted as $\id$ or $e$
    \item Two rotations ($A\to B\to C\to A$ and $A\to C\to B\to A$)
    \item Three reflections % A <-> B
    $A \leftrightarrow B$, $A \leftrightarrow C$, $B \leftrightarrow C$
\end{itemize}
Note that these symmetries preserve the structure of the triangle, hence the composition of two symmetries must also be a symmetry. Let
\begin{itemize}
    \item $\rho$ be the rotation $A\to B\to C\to A$
    \item $\sigma$ be the reflections $B\leftrightarrow C$
\end{itemize}
Note that $\rho\sigma$ is the $A\leftrightarrow C$ reflection and $\sigma\rho$ is the $A\leftrightarrow B$ reflection. Hence they may not be commutative.

% Note that a square has eight symmetries. Then, that tells us that the square and triangle are two different shapes (this applies in general).

We also know that all symmetries can be reversed. $\alpha$ has an inverse $\alpha\inv$ such that $\alpha\alpha\inv = \alpha\inv\alpha = e$. These inspires the following definition:

\begin{definition}
    A \textbf{group} is a set $G$ with a composition
    \begin{align}
        G\times G &\to G \\
        (g, h) &\mapsto g\cdot h
    \end{align}
    Satisfying:
    \begin{itemize}
        \item Associativity: $(g\cdot h)\cdot k = g\cdot (h\cdot k)$
        \item Identity: $\exists\, e\in G$ such that $g\cdot e = e\cdot g = g$ for all $g\in G$
        \item Inverse: $\forall\, g\in G$, $\exists\, g\inv\in G$ such that $g\cdot g\inv = g\inv\cdot g = e$
    \end{itemize}
\end{definition}

Examples:
\begin{itemize}
    \item $\Z$ with $+$ is a group. It is associative, $e = 0$ and $g\inv = -g$.
    \item $\Z/n\Z$ with addition modulo $n$.
    \item If $F$ is a field, it implicitly has two group structures:
    \begin{itemize}
        \item Additive group: $(F,+)$ is a group. It is associative, $e = 0$ and $g\inv = -g$.
        \item Multiplicative group: $(F\setminus\{0\},\times)$ is a group. It is associative, $e = 1$ and $g\inv = 1/g$.
    \end{itemize}
    \item $GL(n,F)$ -- ``general linear group'' contains all invertiable $n\times n$ matrices.
    \item $SL(n,F)$ -- ``special linear group'' contains all invertiable $n\times n$ matrices with determinant $1$.
    \item $SO(n,F)$ -- ``special orthogonal group'' $=\{A\in SL(n,F)|A^t=A\inv\}$.
\end{itemize}
\subsection{Cyclic Groups}
One of the simplest groups is $\Z/n\Z$ for $n\in\N$ with the operation addition modulo n. This is known as the ``cyclic group of order $n$'' or $C_n$. i.e. for $n=8$, $5+7=4\,(\mathrm{mod}\,8)$, which we denote $\bar{5}+\bar{7}=\bar{4}$.

We know the inverse $\bar k\inv=\overline{n-k}$ for nonzero $k$ or $\bar 0\inv=\bar 0.$

Another way to express the cyclic group is $\bar k\leftrightarrow e^{2\pi i k/n}$ with multiplication operation. Then, \begin{align}
    \overline{k+n}=e^{2\pi i (k+n)/n}=e^{2\pi i k/n}e^{2\pi i n/n}=e^{2\pi i k/n}=\bar k.
\end{align}

% draw a circle and 8 equally spaced points on it
\begin{center}
    \begin{tikzpicture}
        \draw (0,0) circle (2cm);
        \foreach \x in {0,1,...,7}
        {
            % draw dot
            \draw (\x*45:2cm) circle (1pt);
        }
        % label the first point
        \node[below] at (0*45:2cm) {$1\leftrightarrow\bar 0$};
        % draw xy axis
        \draw (-3cm,0) -- (3cm,0);
        \draw (0,-3cm) -- (0,3cm);
    \end{tikzpicture}
\end{center}

\begin{definition}[Order]
    The \textbf{order} of a group $G$ is its cardinality denoted $\mathrm{ord}(G)$ or $|G|$. It could be a finite or infinite ordinal. In particular, $|C_n| = n$.
\end{definition}

\subsection{Quaternion Group}
The quaternion group $\mathbb{H}=\{\pm1,\pm i,\pm j,\pm k\}$ is a group of order $8$ with the multiplication operation. It has 

\begin{definition}[Subgroup]
    A \textbf{subgroup} of a group $G$ is a subset $H\subseteq G$ such that $H$ is a group. 
\end{definition}

\begin{definition}[Coset]
    If $G$ is a group and $H\leq G$, consider sets of the form
    \begin{equation}
        Hg=\{hg|h\in H\}
    \end{equation}
    This is a \textbf{right coset} of $H$.
\end{definition}
\begin{theorem}[Partitioning with Cosets]
    Consider $Hg$ and $Hg'$ for $g.g'\in G$. There are two cases:
    \begin{itemize}
        \item They might be disjoint: $Hg\cap Hg'=\emptyset$.
        \item They might intersect. Suppose $hg=h'g'$ for some $h,h'\in H$
        \begin{align}
            h\inv hg&=h\inv h'g'\\
            g&=h\inv h'g'\in Hg'
        \end{align}
        Similarly, $g'\in Hg$. Consider an arbitrary element of $kg\in Hg$ with $k\in H$. Then, $kg=kh\inv h'g'\in Hg'$ i.e. $Hg\leq Hg'$. Similarly, $Hg'\leq Hg$. Thus, $Hg=Hg'$.
    \end{itemize}
    The right cosets of $H$ partition $G$. In particular,
    \begin{equation}
        G = \bigsqcup Hg_i
    \end{equation}
\end{theorem}

For fixed $g$, if $hg=h'g$ for $h,h'\in H$ then $hgg\inv=h'gg\inv$ so $h=h'$. So in $Hg$, every element can be matched with an element of $H$. So, $|Hg|=|H|$.

\begin{theorem}[Lagrange]
    If $|G|<\infty$ and $H\leq G$, then $|H|\Big\vert|G|$
\end{theorem}
\begin{definition}[Index]
    For $H\leq G$, the \textbf{index} of $H$ in $G$ is $[G:H]=|G|/|H|$.
\end{definition}
If $|G|=13$, the only subgroups or $G$ are $\{e\},G$.

If $G=\Z$ and $H=2\Z$ (even numbers). Then $H+0=H$ is one coset, and $H+1=$ the odd integers is another coset. So, $\Z=(2\Z)\sqcup(2\Z+1)$.

\underline{Same for Left Cosets}
Interaction of left and right cosets? 

Consider the triangle group with rotations $e,\rho,\rho^2$ and reflections $\sigma_A, \sigma_B, \sigma_C$ Consider the subgroup $H=\{e,\sigma_A\}$.
\begin{align}
    He&=\{e,\sigma_A\}\\
    H\rho&=\{\rho,\sigma_B\}\\
    H\rho^2&=\{\rho^2,\sigma_C\}\\
    eH=\{e,\sigma_A\}\\
    \rho H&=\{\rho, \sigma_C\}\\
    \rho^2 H&=\{\rho^2,\sigma_B\}
\end{align}
Note that the left and right cosets are different. They are the same if the group is commutative.

\begin{definition}[Action]
    An \textbf{action} of a group $G$ on a set $X$ is a map \begin{align}
        G&\times X\to X \\
        (g,x)&\mapsto gx
    \end{align}
    such that 
    \begin{align}
        (gh)x&=g(hx)\\
        ex&=x
    \end{align}
\end{definition}

If $G$ is a group, it acts on itself. This is called a ``left translation'' or ``left regular action''.

How about the right action $(g,x)\mapsto xg$. The second condition may not be true\begin{align}
    (gh,x)=xgh\\
    (g,(hx))=(g,xh)=xhg
\end{align} which is not true. Instead, let $(g,x)=xg\inv$. Then,
\begin{align}
    (gh,x)=x(gh)\inv=xh\inv g\inv\\
    (g,(h,x))=(g,xh\inv)=xh\inv g\inv
\end{align}
This is the definition of the right action.

There is a third action of $G$ on itself by $(g,x)=gxg\inv.$ This action is called conjugation.

Take the following example: Let $G=SO(3)$ and let $X=S^2$. $G$ acts on $X$ by rotation. Let $H=\left\{\begin{pmatrix}
    \cos\theta & -\sin\theta & 0\\
    \sin\theta & \cos\theta & 0\\
    0 & 0 & 1
\end{pmatrix}\right\}$ be the subgroup of rotations that fixes the $z$-axis. 

$H$ also acts on $X$ ??
\begin{definition}[Orbit]
    If $G$ acts on $X$, the \textbf{orbit} of $x\in X$ is the set $Gx=\{gx|g\in G\}$. i.e. the set of all points $x$ is taken to by elements of $G$.
\end{definition}

The orbits of $H\approx SO(2)$ on the sphere are the lines of latitude (and the north and south poles). 

$H$ fixes the north pole, thus every coset $gH$ takes the north pole to a point. Suppose $gH$ and $g'H$ are cosets such that $gHN=g'HN\implies gN=g'N\implies (g')\inv gN=N\implies (g')\inv g\in H\implies gH...$ so the points ofn the sphere are in 1-1 correspondence with the left cosets of $H$. 
\begin{definition}[Stabilizer]
    If $G$ acts on $X$ and $x\in X$, the ``stabilizer'' of $x$ in $G$ is $\{g\in G|gx=x\}$
\end{definition}

\begin{definition}[Centralizer]
    If $A\subset G$, the \textbf{centralizer} of $A$ in $G$ is $C_G(A)=\{g\in G|ga= ag\forall a\in A\}$
\end{definition}
\begin{itemize}
    \item If $G$ is abelian, then $C_G(A)=G$ for any $A$. 
    \item In the triangle group, $C_G(\{\rho\})=\{e,\rho,\rho^2\}$
\end{itemize}
\begin{definition}[Center]
    The \textbf{center} of $G$ is $Z(G)=\{g\in G|gg'=g'g\forall g'\in G\}=C_G(G)$ 
\end{definition}
\begin{proposition}
    For any $A\subset G$, $C_G(A)\leq Z(G)$ (is a subgroup).
\end{proposition}

Consider the regular $n$-gon ($n\geq3$), what are its rigid motion symmetries?
\begin{itemize}
    \item There are always $n$ rotations by $\frac{2\pi}{n}$ about the origin.
    \item When $n$ is even, there are $n/2$ reflections in each pair of edges, and each pair of vertices. When $n$ is odd, there are $n$ reflections in each pair of (edge, vertex). There are always $n$ reflections.
    \item Write $\rho$ for clockwise rotation by $\frac{2\pi}{n}$. Fix one vertex and let $\sigma$ be the reflection that fixes that vertex.
    \item Note that $\rho\sigma=\sigma\rho\inv$. To show this, it suffices to find where two of the vertices gets mapped.
\end{itemize}
\begin{proposition}
    The symmetries are $e,\rho,\rho^2,\dots,\rho^{n-1},\sigma,\sigma\rho,\sigma\rho^2,\dots,\sigma\rho^{n-1}$
\end{proposition}
\begin{definition}[dihedral group]
    The group of symmetries of the regular $n$-gon is $D_{2n}$, the \textbf{dihedral group} of order $2n$.
\end{definition}

Given $H\leq G$ we write $G/H$ as the set of left cosets
\begin{align}
    G/H&=\{gH|g\in G\}\\
    H\setminus G&=\{Hg|g\in G\}
\end{align}
Both of these are called ``$G$ mod $H$''. In general, the two are \underline{different}. 

Now we want to ask, is $H\setminus G$ a group?
\begin{itemize}
    \item The most naive idea is to reuse multiplication in $G$, i.e. $Hg\cdot Hg'=Hgg'$, but it only sometimes works.
    \item This formula means: $hg\cdot h'g'=h''gg'$. For any $h,h'\in H,\exists h''$ s.t. this holds.
    \item Trick: $hg\cdot h'g'=hgh'eg'=hgh'(g\inv g)g'=h(ghg\inv)gg'$. Now we can ask if $ghg\inv\in H$ (for every $h'\in H$)
\end{itemize}
\begin{definition}[Normal Subgroup]
    A subgroup $H\leq G$ is \textbf{normal} if $ghg\inv\in H\forall g\in G,h\in H$, which is abbreviated as $gHg\inv=H$. $H\trianglelefteq G$ means $H$ is a normal subgroup of $G$
\end{definition}
\begin{itemize}
    \item Notice that if $gHg\inv=H$ then $gH=Hg$. So $H$ is normal, the left and right cosets must be the same.
\end{itemize}

\begin{definition}[Quotient Group]
    If $H\trianglelefteq G$, then $G/H$ is called the quotient group.
\end{definition}
\subsection{Homomorphisms}
\begin{definition}[Homomorphism]
    If $G,K$ are groups, a \textbf{homomorphism} is a map $\varphi:G\to K$ such that $\varphi(gg')=\varphi(g)\varphi(g')\,\forall g,g'\in G$.
\end{definition}
Observations: IF $\varphi:G\to K$ is a homomorphism and $g\in G$, then
\begin{enumerate}
    \item $\varphi(g)=\varphi(eg)=\varphi(e)\varphi(g)$, so $\varphi(e)=e$ (the identity element of $K$)
    \item $e=\varphi(e)=\varphi(gg\inv)=\varphi(g)\varphi(g\inv)$, so $\varphi(g\inv)=\varphi(g)\inv$
\end{enumerate}
Examples
\begin{itemize}
    \item $G=\Z$ and $\varphi:\Z\to\Z, \varphi(n)=2n$ is a homomorphism, as $\varphi(n+m)=2(n+m)=2n+2m=\varphi(n)+\varphi(m)$
    \item $G=\Z,K=\R$ and $\varphi:\Z\to\R,\varphi(n)=n$. This mapping is called an \bluebf{inclusion} as $Z\subset\R$.
    \item If $G$ is a group and $g_0\in G$, then $C_{g_0}:G\to G,g\mapsto g_0gg_0\inv$ is a homomorphism.
    \item A linear transformation $T:V\to W$ if $V,W$ are vector spaces (the additive group). 
    \item Note that $\varphi:g\mapsto g\inv$ is \bluebf{only} a homomorphism if $G$ is abelian.
\end{itemize}
\begin{definition}[Kernel/Image]
    If $\varphi:G\to G'$ is a homomorphism, then the \textbf{kernel} of $\varphi$ is\begin{equation}
        \ker(\varphi)=\{g\in G|\varphi(g)=e\}.
    \end{equation}
    The \textbf{image} of $\varphi$ is\begin{equation}
        \mathrm{im}(\varphi)=\{\varphi(g)|g\in G\}\subseteq G'
    \end{equation}
\end{definition}
\begin{theorem}
    $\ker(\varphi)\leq G$ and $\mathrm{im}(\varphi)\leq G'$ $\ker(\varphi)\trianglelefteq G$
    \begin{proof}
        Since $\varphi(e)=e$, $e\in\ker(\varphi)$, and $e\in\mathrm{im}(\varphi)$.
        So both are nonempty.

        Suppose $g,h\in\ker(\varphi)$, $e=\varphi(e)=\varphi(hh\inv)=\varphi(h)\varphi(h\inv)$ \dots
    \end{proof}
\end{theorem}
\begin{itemize}
    \item Suppose $N\trianglelefteq G$ and then define $G\to G/N, g\mapsto Ng$. We claim this is a homomorphism. Proof is simple $\varphi(gg')=Ngg'$, $\varphi(g)\varphi(g')=NgNg'=NgN(g\inv gg')=N(gNg\inv)gg'=NNgg'=Ngg'$ 
    \item This map is called the (natural) \bluebf{projection} of $G$ onto $G/N$. Sometimes written $\Pi_{G/N}$ or $\mathrm{proj}_{G/N}$.
    \item $\mathrm{im}(\Pi_{G/N})=G/N$ and $\ker(\Pi_{G/N})=N$.
    \item Any homomorphism is related to this one, so this could be considered as the ``generic homomorphism''.
\end{itemize}
\begin{definition}[Isomorphism]
    If $\varphi:G\to H$ is a homomorphism, and $\ker(\varphi)=\{e\}$ then $\varphi$ is injective.
    If $\varphi(G)=H$ then $\varphi$ is surjective. Thinking of $G$ and $H$ as sets, there is an inverse $\varphi\inv:H\to G$ such that $\varphi\inv\circ\varphi=1_G$ and $\varphi\circ\varphi\inv=1_H.$ It is easy to check that $\varphi\inv$ is also a homomorphism.

    In this case, $\varphi$ is an \textbf{isomorphism}
\end{definition}
\begin{itemize}
    \item Suppose we have an injective homomorphism $\varphi:G\to H$ where $\ker(\varphi)=\{e\}.$ Then, we can consider $\varphi:G\to\mathrm{im}(\varphi)<H.$ Sometimes we say $\varphi:G\to H$ is an \bluebf{isomorphism into} $H,$ as opposed to an isomorphism \bluebf{onto} $H$ or between $G$ and $H.$
\end{itemize}
\begin{definition}[Automorphism]
    If $G$ is a group, an \textbf{automorphism} of $G$ is an isomorphism $\varphi:G\to G.$
\end{definition}
Examples:
\begin{itemize}
    \item If $G=\Z,n\mapsto-n$ is the only automorphism apart from the identity.
    \item If $G$ is abelian, $g\mapsto g\inv$ is an automorphism.
    \item If $F$ is a field, and $G=GL(n,F)$ then $g\mapsto(g^t)\inv$ (transposed inverse) is an automorphism.
    \item If we fix $g_0\in G$ then the conjugation $C_{g_0}:G\to G$ where $C_{g_0}(g)=g_0g{g_0}\inv$ is an automorphism. 
\end{itemize}
\begin{definition}[Automorphism Group]
    $\mathrm{Alt}(G)$ is the \textbf{group} of automorphisms of $G$.
\end{definition}
\begin{definition}[Inner/Outer Automorphisms]
    The \textbf{inner automorphisms} of $G$ are \begin{equation}
        \textrm{Inn}(G)=\{\varphi\in\mathrm{Alt}(G)|\varphi=C_{g_0}\text{ for some }g_0\in G\}.
    \end{equation}
    If an element of $\mathrm{Alt}(G)$ that is not inner is \textbf{outer}.
\end{definition}
\begin{itemize}
    \item It is easy to show that $\mathrm{Inn}(G)\leq\mathrm{Alt}(G).$ 
    \item Observe that if $G$ is abelian, then $\mathrm{Inn}(G)=\{\mathrm{id}\}$
    \item In general, $\{\mathrm{id}\}\leq\mathrm{Inn}(G)\leq\mathrm{Alt}(G).$
\end{itemize}
\begin{itemize}
    \item The map \begin{align}
        G&\to\mathrm{Alt}(G)\\
        g&\to C_g
    \end{align}
    is a homomorphism. Its image is $\mathrm{Inn}(G)$ and its kernel is $Z_G$ (the center).
\end{itemize}
\begin{definition}[Fiber]
    If $p$ is a projection, then $p\inv(x)$ is the \textbf{fiber} over $x$
\end{definition}
\begin{itemize}
    \item If $N\triangleleft G,$ the projection $\pi:G\to G/N$ is a homomorphism. The fibers of $\pi$ is the cosets $gN=Ng,$ and they are all the same size.
    \item Suppose $\varphi:G\to H$ is a homomorphism, and $N=\ker(\varphi)\trianglelefteq G.$ The fibers of $\varphi$ is the cosets of $G/N.$
    \item We have $\varphi:G\to H$ and $\pi:G\to G/N.$ Wouldn't it be nice if $G/N\to H$ ``induced by $\varphi$'' were a homomorphism? Well, it is. 
\end{itemize}
\begin{theorem}[(First) Isomorphism]
    If $\varphi:G\to H$ is a homomorphism, and $N=\ker(\varphi),$ then there is a homomorphism $\bar\varphi:G/N\to H$ such that $\bar\varphi\circ\pi=\varphi.$ 
    
    Moreover, $\ker(\bar\varphi)=\{eN\},$ the trivial subgroup of $G/N,$ so $\bar\varphi$ is injective. So, $\bar\varphi:G/N\to\mathrm{im}(\varphi)$ is an \bluebf{isomorphism}.
\end{theorem}
\begin{itemize}
    \item This theorem suggests that you can construct an isomorphism from an arbitrary homomorphism. First, $\varphi$ factors through $G/N,$ then we can include it into $H.$
    \begin{equation}
        G\to^\pi G/N\to^{\bar\varphi}\mathrm{im}(\varphi)\to^{\text{inclusion}}H
    \end{equation}
\end{itemize}

\begin{theorem}[(Third) Isomorphism]
    $N\trianglelefteq G$ and $H\leq G,$ then $N\leq H\implies N\trianglelefteq G.$
    \begin{proof}
        ????
    \end{proof}
\end{theorem}
\begin{theorem}
    \begin{equation}
        G/H\cong G/N\Big/H/N
    \end{equation}
    \begin{proof}
        Define $\varphi:G\to G/N\Big/H/N$ by \begin{align}
            \varphi(g)=(gN)H/N
        \end{align}
        We need to show $\varphi$ is a homomorphism. Let \begin{align}
            \varphi(gg')&=gg'N\:H/N\\
            &=gNg'N\:H/N\\
            &=gN\:H/N\cdot g'N\:H/N\\
            &=\varphi(g)\varphi(g')\\
        \end{align}
    \end{proof}
    We will then ask what is $\ker(\varphi).$ Suppose $\varphi(g)=H/N,$ so $gN\:H/N=H/N.$ But $g$ is a representation for $gN,$ so $gH/N$ for this to be in $H/N$ we want $g\in H$ so $\ker(\varphi)=H.$ An arbitrary element of $G/N\Big/H/N$ is $gN\:H/N$ for some $g\in G,$ so $\mathrm{im}(\varphi)=G/N\Big/H/N.$
\end{theorem}
\begin{itemize}
    \item $G=\Z,H=3\Z,K=4\Z.$ By the second isomorphism theorem, $\Z/3\Z\cong 4\Z/12\Z,$ and also $Z/4\Z\cong 3\Z/12\Z.$
\end{itemize}
\begin{definition}[Equivilence Class]
    Being in the same coset of a subgroup $H$ is an equivalence relation. So, the large group is a disjoint union of equivalence classes (cosets) of $H.$
\end{definition}
\begin{itemize}
    \item The cosets of $\Z$ in $\R$ is $r+\Z$ for $r\in[0,1).$
    \item Homomorphism $\varphi:\R\to\C^\times, t\mapsto e^{2\pi it}.$ Then, $\ker(\varphi)=\Z.$ Observe tat $\varphi$ is \bluebf{onto} the unit circle, by the first isomorphism theorem, $\R/\ker(\varphi)=\R/\Z\cong S^1.$
    \item $\Z^2\triangleleft\R^2$
\end{itemize}
\begin{theorem}[Fourth Isomorphism Theorem/Lattice Theorem]
    Consider a lattice of subgroups with $N\trianglelefteq G.$ In $G/N,$ the subgroup lattice has the same structure as the subgroup lattice of $G$ that contains $N.$

    Specifically, if $N\trianglelefteq G,$ and $N\trianglelefteq H<G,$ we write $\bar H=H/N.$ Including $\bar G=G/N$ and $\bar N=\bar e=N/N.$

    Then, the lattice of $\bar H$s in $\bar G$ has the same lattice structures as the part of the lattice for $G$ consisting of subgroups that are intermediate between $N$ and $G.$ Moreover,\begin{align}
        H\leq K&\iff\bar H\leq\bar K\\
        H\trianglelefteq K&\iff\bar H\trianglelefteq\bar K\\
        [H:K]&=[\bar H:\bar K] \text{ if }K\leq H\\
        \overline{H\cap K}&=\bar H\cap\bar K\\
        \overline{\langle H,K\rangle}&=\langle\bar H,\bar K\rangle
    \end{align}
\end{theorem}
If $G,G'$ are groups, consider the cartesian product $G\times G'=\{(g,g')|g\in G,g'\in G'\}.$ Note that $|G\times G'|=|G||H|.$ There is an obvious way to turn this into a group by\begin{align}
    (g,g')(h,h')&=(gh,g'h')\\
    (g,g')^{-1}&=(g^{-1},g'^{-1})
    e=(e,e)
\end{align}
In $G\times G',$ the subset $G_0=:\{(g,e)|g\in G\}\cong G$ is a subgroup. Likewise, $G'_0=:\{(e,g')|g'\in G'\}\cong G'.$ Also notice that $G_0$ and $G'_0$ commute. So, $(G\times G')/G_0\cong G'.$

\subsection{Symmetric Groups}
\begin{definition}[Symmetric Group]
    The symmetric group $S_n$ is the group of permutation of $n$ elements, with composition as the operation.
\end{definition}
\begin{itemize}
    \item $|S_n|=n!$
    \item A cycle is a permutation that cycles through some subset of $\{1,\dots,n\}$, denoted as \begin{align}
        (a_1\,a_2\,\dots\,a_k),\quad k\leq n\text{ and }a_i\text{ are distinct.}
    \end{align}
    Represents the permutation $a_1\to a_2\to\dots\to a_k\to a_1.$
    \item Note that these are ambiguous, as $(a_1\,a_2\,\dots\,a_k)$ is the same as $(a_2\,a_3\,\dots\,a_k\,a_1).$ So by convention, we often start with the smallest number first so they are unique. 
    \item $k$ is the length of the cycle, it is also called a \bluebf{$k$-cycle}.
    \item Every permutation can be written as a product of disjoint cycles. If given a permutation, we will start from $1$ and write a cycle until we get back to $1.$ Then, we will start from the next number that hasn't been included yet and repeat until we get to the end.
    \item If $\sigma=(1\,3\,6)(4\,5),$ then $\sigma\inv=(4\,5)\inv(1\,3\,6)\inv=(4\,5)(1\,6\,3)=(1\,6\,3)(4\,5).$ We will order the cycles by their first element, and omit $1$-cycles.
    \item Two \bluebf{disjoint cycles} (i.e. without any numbers in common) will commute.
    \item If cycles are not disjoint, like $\sigma=(1\,4\,2)(2\,3\,5)(3\,4\,7)\in S_7$ will not commute.
    \begin{itemize}
        \item $1\to4$
        \item $4\to7$
        \item $7\to3\to5$
        \item $5\to2\to1$
        \item $2\to3$
        \item $3\to4\to2$
    \end{itemize}
    So $\sigma=(1\,4\,7\,5)(2\,3).$
    \item Any $k$-cycle is a product of $2$-cycles. Thus, every element in the symmetric group can be written as a product of $2$-cycles so $S_n$ is generated by $2$-cycles. For example, if $k=4$ and $\sigma=(a\,b\,c\,d),$ then $\sigma=(a\,d)(a\,c)(a\,b).$
    \item We can ask what is the minimum number of $2$-cycles needed to generate any $\sigma\in S_n.$ In general, this is a very difficult question to answer. However, the \bluebf{parity} of the number of $2$-cycles in a product equalling $\sigma$ is well-defined.
    \item If $\sigma=(a_1\,b_1)(a_2\,b_2)\dots(a_k\,b_k)$ is a product of $2$-cycles, then $\sigma$ is \bluebf{even} if $k$ is even, and \bluebf{odd} if $k$ is odd.
    \item \redbf{Warning: a $k$-cycle is even if $k$ is odd, and odd if $k$ is even.}
    \item To make odd and even well defined, we need to know that the parity of a permutation is independent of the way we write the cycles.
    \begin{proof}
        Given $\sigma\in S_n$ is a $k$-cycle. Define $\Delta=\prod_{1\leq i<j\leq n}(j-i).$ If $\tau\in S_n$, it acts on $\Delta$ with \begin{align}
            \tau\cdot\Delta=\prod_{1\leq i<j\leq n}(\tau(j)-\tau(i)).
        \end{align}
        These two products are the same up to a factor of $\pm1,$ you have to multiply by $-1$ for each pair $i<j$ for which $\tau(i)>\tau(j).$
        
        We will consider how $(a\,b)$ with $a<b$ affect $\Delta.$ If neither $i$ nor $j$ is equal to $a$ or $b,$ the term is unaffected. Note that \begin{itemize}
            \item If $i<a,$ then $i<\tau(a)=b$ and $i<\tau(b)=a.$ So $(i\,a)$ or $(i\,b)$ are unaffected.
            \item Likewise, for $j>b$ then $(a\,j)$ or $(b\,j)$ are unaffected.
        \end{itemize}
        The only pairs that will be affected are ones $(a\,i),(i\,b)$ with $a<i<b$ and $(a\,b).$ If $a<i<b,$ then both $(a\,i)$ and $(i\,b)$ will change sign, so the product will be unaffected. $(a\,b)$ will change sign, so $\Delta$ will change sign under a transposition.

        If $\sigma\in S_n,$ write it as any product of $k$ transpositions. If $\sigma\cdot\Delta=\Delta$ then there must be an even number of transpositions. If $\sigma\cdot\Delta=-\Delta$ then there must be an odd number of transpositions. Thus, the parity of $\sigma$ is independent of the way we write it.
    \end{proof}
\end{itemize}
\begin{definition}[Sign]
    The sign of $\sigma\in S_n$ is \begin{equation}
        \sgn(\sigma)=(-1)^k,
    \end{equation}
    if $\sigma$ is a product of $k$ transpositions. 
\end{definition}
\begin{itemize}
    \item Note that $\sgn(\sigma\tau)=(-1)^k(-1)^l=(-1)^{k+l}=\sgn(\sigma)\sgn(\tau).$
    \item Thus, $\sgn:S_n\to\{\pm1\}$ is a homomorphism.
    \item $\ker(\sgn)=A_n\trianglelefteq S_n$ is the alternating group of $k$ elements which contains all the even permutations. Note that
    \begin{align}
        S_n/A_n\cong\{\pm1\}\quad[S_n:A_n]=2\quad|A_n|=\frac{n!}{2}
    \end{align}
    \item for $n>5,$ $A_n$ has no normal subgroups. What are the possible cycle types in $A_5$? There is $(a\,b\,c\,d\,e),(a\,b)(c\,d),(a\,b\,c)$
    \item Let $\sigma\in S_n$ with $a\to b\to c\to\cdots,$ and suppose $\tau\in S_n$ takes $a\to a',b\to b',c\to c',\dots.$ Consider the conjugation $\tau\sigma\tau\inv.$\begin{align}
        \tau\sigma\tau\inv(a')=\tau\sigma(a)=\tau(b)=b'\\
        \tau\sigma\tau\inv(b')=\tau\sigma(b)=\tau(c)=c'\\
    \end{align}
    So $\tau\sigma\tau\inv$ takes $a'\to b'\to c'\to\cdots.$ Conjugating by $\tau$ ``relabels'' what $\sigma$ by replacing $a$ with $a',$ \dots.
\end{itemize}
\subsection{Simple Group}
One way we study groups is to write it as a chain of normal subgroups $G_0=\{e\}\triangleleft G_1\triangleleft G_2\triangleleft\cdots\triangleleft G_r=G,$ where $G_{i+1}/G$ is a simple group $\forall i=0,\dots,r-1.$ A decomposition like this is called a \bluebf{Jordan-Holder Series} (composition series), and the quotients are called the \bluebf{composition factors}. However, the same $G$ may have different composition series. \begin{theorem}[Jordan-Holder]
    Any two Jordan-Holder series for $G$ have the same length. Moreover, the composition factors are the same (but perhaps in different orders).
\end{theorem}

Example: Suppose $H,K$ are both normal subgroups of $G.$ Apply 2nd isomorphism theorem. Note, $H\subseteq Norm_G(K)=G$ and $K\subseteq Norm_G(H)=G.$ Thus, $HK/K\cong H/H\cap K$ and $HK/H\cong K/H\cap K.$ In this example there are two composition series\begin{align}
    \{e\}\triangleleft H\cap K\triangleleft H\triangleleft HK\triangleleft G\\
    \{e\}\triangleleft H\cap K\triangleleft K\triangleleft HK\triangleleft G
\end{align}

Every group has a Jordan-Holder series. In genera, a group $G$ is not determined by its Jordan-Holder series. However, if $G$ is simple, then its Jordan-Holder series is $\{e\}\triangleleft G.$

\begin{definition}[Solvable]
    If the composition factor $G_{i+1}/G_i$ of $G$ are all \textbf{abelian}, we say $G$ is \bluebf{solvable}.
\end{definition}

If $G$ acts on a set $X,$ then each $g\in G$ permutes the element of $X.$ So there is a map $G\to S_X$ (the symmetric group of $X$). It is easy to show that this map is a homomorphism. So, we will allow ourselves to go between group actions and Homomorphisms into $S_X.$

Suppose $H\leq G$ and let $X=G/H$ be the coset space. So, $G$ acts on $X$ by left multiplication $g(xH)\mapsto gxH.$ If $n=[G:H]=|X|,$ the action amounts to a homomorphism $\varphi:G\to S_n.$

Our first observation is that $G$ acts \textbf{transitively}. For any $x,y\in X,\,\exists\,g\in G$ s.t. $gx=y.$ i.e. the orbit of any $x\in X$ is $X.$

What is $\ker\varphi?$ We know that if $h\in\ker\varphi,$ that $hxH=xH.$ Then consider $h',h''\in H$ then \begin{align}
    hxh'=xh''\\
    hx=xh''{h'}\inv\\
    h=xh''{h'}\inv x\inv\\
    \ker\varphi=\bigcap_{x\in G}xHx\inv
\end{align}
If $H=\{e\},$ then $G/H=G,$ so $\ker\varphi=\{e\}.$ then $\varphi$ is injective. By the first isomorphism theorem, $G\cong \mathrm{im}\,\varphi=S_n.$
\begin{theorem}[Cayley]
    Any group $G$ with $|G|=n$ is isomorphic to a subgroup of $S_n.$
    \begin{proof}
        We already proved it!
    \end{proof}
\end{theorem}
Another example is to let $G$ act on itself by conjugation. In this case, $\varphi$ with $g\cdot x=gxg\inv=C_g(x).$ This is not a transitive action unless $G$ is trivial.  The orbits of conjugation are the \bluebf{conjugacy classes} of $G.$ They are disjoint (because conjugacy is an equivalence relation). 

Note that $geg\inv=e,\,\forall g.$ If $z\in Z(G),$ then $gzt\inv=zgg\inv=z\,\forall g,$ then the conjugacy classes contain a single element.

If $G$ is abelian, $Z(G)=G$ and every element is its own conjugacy class.

Because conjugacy is an equivalence relation, $G$ is a disjoint union of all conjugacy classes.

If $Z(G)=\{e,z_1,\dots,z_k\}$ and $g_1,\dots,g_m$ are representatives from the non-central conjugacy classes. Let's write $C(g_i)=\{gg_ig\inv|g\in G\}.$ So,\begin{align}
    G=Z(G)\sqcup\left(\bigsqcup C(g_i)\right)
\end{align}
so \begin{align}
    |G|=|Z(G)|+\sum_i|C(g_i)|
\end{align}
This is called the \bluebf{Class Equation}.\begin{theorem}[Orbit-Stabilizer]
    If $G$ acts on $X,$ for each $x\in X,$ write $G\cdot x$ for its orbit. Then,\begin{align}
        |G\cdot x|=[G:G_x]=[G:\mathrm{Stab}(x)]
    \end{align}
    The point is that two things in the same coset of $G_x$ has the same effect on $x$.
\end{theorem}
Under conjugation, \begin{align}
    \mathrm{Stab}(x)=G_x=\{g\in G|gxg\inv=x\}=Z(x),
\end{align}
the centralizer of $x$. So the class equation can be rewritten as \begin{align}
    |G|=|Z(G)|+\sum_{i}[G:Z(g_i)]
\end{align}
\begin{definition}[$p$-group]
    Suppose $p$ is prime, $G$ is a \bluebf{$p$-group} if $|G|=p^k$ for some $k\geq 1.$
\end{definition}
\begin{theorem}
    If $G$ is a non-trivial $p$-group, then it has a non-trivial center.
    \begin{proof}
        Suppose $|G|=1.$ Then \begin{align}
            |G|=|Z(G)|+\sum_{i}[G:Z(g_i)]
        \end{align}
        Claim $Z(g_i)<G$, otherwise $g_i\in Z(G).$ By Lagrange's theorem $|Z(g_i)|\mid|G|=p^k.$ So $|Z(g_i)|=p^l$ for some $l<k.$ Then,\begin{align}
            p^k=|G|=|Z|+\sum_{i}[G:Z(g_i)]\\
        \end{align} 
        Since $|Z|=1,$ the RHS is not divisible by $p$ so this is a contradiction.
    \end{proof}
    \begin{corollary}
        Suppose $p$ is prime. If $|G|=p^2,$ then $G$ is abelian.
        \begin{proof}[Proof]
            We know $Z(G)$ is a non-trivial subgroup so $1\neq|Z(G)|\Big|p^2.$ So $|Z(G)|=p$ or $p^2.$ If $|Z(G)|=p^2,$ then $G$ is abelian by definition. If $|Z(G)|=p,$ then $|G/Z(G)|=p$ hence $G/Z(G)\cong C_p.$ So $x\notin Z(G),$ then $G/Z(G)=\{\bar e, \bar x, \bar x^2,\dots\bar x^{p-1}\}$ where $\bar x=:xZ(G).$ Also, $\bar x^p=\bar e\in G/Z(G).$ Note that $\mathrm{ord}(x)$ is either $p$ or $p^2$.
            \begin{itemize}
                \item If $|\langle x\rangle|=p^2$ so $\langle x\rangle=G$ and $G$ is cyclic hence abelian. 
                \item If $\mathrm{ord}(p),$ then $G=\bigcup_{k=0}^{p-1}x^kZ(G).$ Recall $|Z(G)|=p,$ so $Z(G)$ is cyclic. Then, \begin{align}
                    Z(G)=\{e,z,z^2,\dots,z^{p-1}\}
                \end{align}
                so \begin{align}
                    G=\{x^iz^j|0\leq i,j<p\}
                \end{align}
                These elements commute. $x^iz^jx^mz^n=x^ix^mz^jz^n=x^{i+m}z^{j+n}$
            \end{itemize}
        \end{proof}
    \end{corollary}
\end{theorem}
Note we need to be careful with the steps in this proof. Just because $\bar x^p=\bar e$ doesn't mean there is a representative $x\in\bar x$ that is order $p.$

\begin{itemize}
    \item Now we consider the rotations of a tetrahedron. A easy way to think about this is to identify a ``top'' vertex, which is well defined ($4$ possibilities). Then, we fix the top and we have $3$ rotations (like of the triangle). So, there are $12$ rotations.
    \item Apart from $e$, there are two non-trivial rotations that fix any particular vertex. This only accounts for $8$ rotations, and $e$, so we are missing $3$ rotations. 
    \item The other rotations does not fix any vertices and are like $(1\,2)(3\,4).$ Then, $2,3,4$ goes with $1$ so we have $3$ rotations. This accounts for all 12.
    \item In summary, we have $e,$ and $8$ rotations in the form $(a\,b\,c)$ and $3$ rotations in the form $(a\,b)(c\,d).$ This is $A_4.$
    \item The rigid motions are $S_4.$
\end{itemize}
\begin{proposition}
    $A_5$ is simple. $A_5\triangleleft S_5$ with index $2,$ so $|A_5|=60.$
    
    \begin{proof}
        We will enumerate the conjugacy classes of $S_5$ \begin{itemize}
            \item $(a\,b\,c\,d\,e)\in A_5$
            \item $(a\,b\,c\,d)\notin A_5$
            \item $(a\,b\,c)\in A_5$
            \item $(a\,b\,c)(d\,e)\notin A_5$
            \item $(a\,b)(c\,d)\in A_5$
            \item $(a\,b)\notin A_5$
            \item $e\in A_5$
        \end{itemize}
        There are $24$ elements in the conjugacy class of $(a\,b\,c\,d\,e).$ However, $24$ does not divide $60$ so it is not a conjugacy class of $A_5.$

        Consider the centralizer $Z_{A_5}(a\,b\,c\,d\,e)\geq\left\langle(a\,b\,c\,d\,e)\right\rangle$ which has order $5$. But $Z_{A_5}(a\,b\,c\,d\,e)\leq Z_{S_5}(a\,b\,c\,d\,e)$ so $Z_{A_5}(a\,b\,c\,d\,e)=\left\langle(a\,b\,c\,d\,e)\right\rangle$

        So there are two $A_5$ conjugate classes of $5$-cycles, each with $12$ elements.

        There are $20$ $3$-cycles in $S_5.$ Are they all conjugate in $A_5?$ If $(a\,b\,c)$ is conjugate to $(x\,y\,z)$ by $\sigma\in S_5$, then it is also conjugate by $\sigma(d\,e)$ If $\sigma\notin A_5$ then $\sigma(d\,e)\in A_5$ so there is one conjugate class of $20$ $3$-cycles. 

        There are $15$ double transpositions.

        \textbf{If we have a normal subgroup, it is a union of the conjugacy classes,} so if $A_5$ has a normal subgroup it must be a combination of $1+15,20,12,12$ but there is no combination (apart from $1$) that divides $60.$ Hence, $A_5$ is simple. 
    \end{proof}
\end{proposition}
\section{Sylow Theorems}
\begin{theorem}
    Suppose $|G|=p^\alpha n$ where $p\not|n.$ Then, a subgroup $P\leq G$ is a Sylow $p$-subgroup if $|P|=p^\alpha.$ We'll write $n_p(G)$ for the number of Sylow $p$-subgroups of $G.$ \begin{enumerate}
        \item Sylow $p$-subgroups exist.
        \item Suppose $P$ is a Sylow $p$-subgroup of $G$ and $Q\leq G$ s.t. $|Q|=p^r$ for some $r>0.$ Then, $\exists\,g\in G$ s.t. $gQg\inv\subseteq P.$ In particular, all Sylow $p$-subgroups of $G$ are conjugate. 
        \item $n_p(G)\equiv1\mod{p}$ and $n_p(G)=[G:\mathrm{Norm}_G(P)]$ for any Sylow $p$-subgroup. Hence $n_p(G)||G|.$ It actually also divides $n=|G|/|P|.$
    \end{enumerate}
\end{theorem}
Before proving the theorem we will consider the following example: Let $G=S_3.$ The Sylow $2$ subgroups are $\{e,(1\,2)\},\{e,(1\,3)\},\{e,(2\,3)\}$ we know $n_2(S_3)=3\equiv1\mod2.$ The only Sylow $3$ subgroup is $A_3,$ so $n_3(S_3)=1.$\begin{lemma}
    If $G$ is abelian and $p\Big||G|,$ then $G$ contains an element of order $p.$
    \begin{proof}
        If $|G|=p$ then $G$ is cyclic and every non-trivial element has order $p$.

        If $|G|>p,$ and $x\in G$ with order $p^rm$ where $p\not{\big|}m.$ If $r\neq0,$then $x^{p^{r-1}m}$ has order $p.$
        
        This reduces us to the case where $p\nmid\mathrm{ord}(x), \forall x\in G.$ We will use induction. \begin{itemize}
            \item Assume the result is true for all groups smaller than $G.$ 
            \item If $p\nmid\mathrm{ord}(x)=|\langle x\rangle|<|G|.$ As $G$ is abelian, then $N=:\langle x\rangle\triangleleft G.$ 
            \item By induction $G/N$ contains an element of order $p.$
            \item i.e. $\exists y=y_0N\in G/N$ s.t. $y^p=e=N.$ so $y_0^p\in N$
            \item We claim that $\langle y_0^p\rangle<\langle y_0\rangle$ since otherwise $y_0\in N$ which has order $1.$
            \item This means $p\mid|y_0|$ otherwise $\langle y_0^p\rangle=\langle y_0\rangle.$ This is a contradiction.
            \item This means a suitable power of $y_0$ must have order $p.$
        \end{itemize}
    \end{proof}
\end{lemma}
\begin{lemma}
    If $P\in\mathrm{Syl}_p(G)$ and $Q$ is a non-trivial $p$-subgroup of $G.$ Then, $Q\cap\mathrm{Norm}_G(P)=Q\cap P.$
    \begin{proof}
        Let $H=Q\cap\mathrm{Norm}_G(P)\geq Q\cap P.$ We need to show that $H\leq Q\cap P.$
        But $H\leq Q$ so we only need to show that $H\leq P.$

        $H\leq N_G(P)\implies HP$ is a subgroup. The result will follow if we can argue that $HP$ is a $p$-group. We know that \begin{align}
            |HP|=\frac{|H||P|}{|H\cap P|}
        \end{align}
        Since $|H|,|P|,|H\cap P|$ are all powers of $P.$ So $HP\geq P$ but $|HP|$ can't be bigger than $|P|.$
    \end{proof}
\end{lemma}

Proof that Sylow $p$-subgroup exists. We will use induction on $|G|.$

If $p\mid |Z(G)|,$ we know by the lemma that $\exists z\in Z(G)$ with $|z|=p.$ Let $N=\langle z\rangle$ is a normal subgroup as $N\leq Z(G).$ Then, $G/N$ is a smaller group than $G.$ By the induction hypothesis say $G/N$ has a Sylow $p$-subgroup.

If $|G|=p^\alpha m,p\nmid m$ then $G/N=p^{\alpha-1}m.$ So, it has a Sylow $p$-subgroup of order $p^{\alpha-1}.$ By the lattice isomorphism theorem, the preimage of this group in $G$ has order $p^\alpha,$ as required.

Assume $p\nmid|Z(G)|.$ Let $g_1,\dots,g_k$ be representatives of the non-central conjugacy classes of $G.$ So,\begin{align}
    |Z(G)|+\sum_{i=1}^k[G:C_G(g_i)]=|G|.
\end{align}
We know that $p\mid |G|$ but $p\nmid |Z(G)|$ meaning for some $i,$ we know $p\nmid[G:C_G(g_i)].$ As $g_i$ represents a non-central conjugacy class, then $C_G(g_i)<G.$ We will use the induction hypothesis. Note that since $p\nmid[G:C_G(g_i)],$ then $p^\alpha\mid|C_G(g_i)|.$ By the induction hypothesis, $C_G(g_i)$ has a Sylow $p$-subgroup of order $p^\alpha,$ it is also a Sylow $p$-subgroup of $G.$ Thus, Sylow $p$-subgroups exist.

Fix a Sylow $p$-subgroup $P_1$ of $G$ and enumerate all its distinct conjugates as $P_1,\dots,P_r.$ Let $Q$ be any $p$-subgroup.

$G$ acts on $\mathcal S=\{P_1,\dots,P_r\},$ and $Q$ also act on $\mathcal S,$ but it may not have a single orbit. Decompose $\mathcal S$ into $Q$ orbits, \begin{align}
    S=\mathcal O_1\sqcup\mathcal O_2\sqcup\cdots\sqcup\mathcal O_s.
\end{align}
How big is $\mathcal O_k$? Relabel $P_1,\dots P_k$ s.t. $\mathcal O_k=\{qP_kq\inv|q\in Q\}.$ We know how to find the size of a conjugacy class. $|\mathcal O_k|=[Q:N_Q(P_k)].$ Note that $N_Q(P_k)=N_G(P_k)\cap Q.$ The second lemma states $N_G(P_k)\cap Q=P_k\cap Q.$

For now, let $Q=P_1.$ So, $\mathcal O_1=\{qP_1q\inv|q\in P_1\}=P_1.$ \begin{align}
    |S|=r=\underbrace{|\mathcal O_1|}_{1}+\underbrace{\sum_{i=2}^s[P_1:P_1\cap P_i]}_{\text{divisible by }p}
\end{align}
If we know that if all Sylow $p$-subgroup are conjugate, then we know the number of Sylow $p$-subgroup is $1\mod p.$

Let $Q$ be any $p$-subgroup of $G$ and suppose $Q$ is not contained in any of the $P_1,\dots P_r.$ Then, $Q\cap P_i$ is a proper subgroup of $Q.$ Then, \begin{align}
    |\mathcal O_k|=[Q:P_k\cap Q]
\end{align}
is divisible by $P.$ So, $p\mid|\mathcal S|$ so $p\mid|r$ but $r\equiv 1\mod p$ so this is a contradiction. 


Suppose $|G|=pq,$ and $p<q$ prime. We know $n_q(G)=1,$ i.e. $\mathrm{Syl}(G)=\{Q\},$ then $Q\triangleleft G.$ One possibilities is that $G$ is cyclic. Often, $n_p(G)=0$ unless $p\mid(q-1),$ then it is more complicated.

The significance os $q-1=$ the number of units $\mod q,$ which turns out to be the number of automorphisms of $C_q.$ (multiply each element of $C_q$ by a unit $u=(\Z/q\Z)^x.$

So this is a homomorphism $C_p\to\mathrm{Aut}(C_q).$ We can use this homomorphism to make a group that is not abelian.

\begin{definition}[Finitely Generated]
    An abelian group $G$ is \bluebf{finitely generated} if there exists a finite set $S$ such that $G=\langle S\rangle.$
\end{definition}
Examples of finitely generated abelian groups are finite abelian groups, $\Z,\Z^r$ but not $\R,S^1,\mathbb Q$.
\begin{theorem}[Fundamental Theorem of Abelian Groups]
    If $G$ is a finitely generated abelian group, then $G$ is isomorphic to a product \begin{align}
        G\cong\Z^r\times\prod_{i=1}^sC_{r_i}
    \end{align}
    where $r\in\N_0,n_i\in\N^{>1},$ and $n_{i+1}\mid n_i\forall i=1,\dots,s-1.$ Note the following:\begin{itemize}
        \item $r=0\iff G$ is finite.
        \item $G$ is cyclic $\iff r=0\land s=1$
    \end{itemize}
    Moreover, this decomposition is unique up to isomorphism.
    \begin{proof}
        The proof will come easily from another theorem later.
    \end{proof}
\end{theorem}
\begin{definition}
    In this decomposition, $r$ is called the \bluebf{free rank} of $G$ or the \bluebf{Betti number}Other Information of $G.$ The $n_i$'s are called the \bluebf{invariant factors} of $G.$
\end{definition}

Another version. Any finitely generated abelian group $G$ can be written as \begin{align}
    G=\Z^r\times\prod_{i=1}^kP_{p_i}
\end{align}
where $|G/\Z^r|=\prod_i p_i^{\alpha_i}.$ Moreover, for each $i,$ \begin{align}
    P_{p_i}=C_{p_i^{\beta_1^i}}\times C_{p_i^{\beta_2^i}}\times\cdots\times C_{p_i^{\beta_{\alpha_i}^i}}
\end{align}
where $\beta_1^i\geq \beta_2^i\geq\cdots\geq\beta_{\alpha_i}^i$ and $\beta_1^i+\beta_2^i+\cdots+\beta_{\alpha_i}^i=\alpha_i.$
The notation is awful, but idea is we can decompose $G$ into its Sylow $p$-subgroups and then decompose each Sylow $p$-subgroup into its cyclic factors. This decomposition is unique up to isomorphism.

\begin{definition}[Elementary Divisors]
    The subgroups $C_{p_i^{\beta_1^i}},\dots,C_{p_i^{\beta_{\alpha_i}^i}}$ or sometimes their orders are called the \bluebf{elementary divisors} of $G.$
\end{definition}

Semidirect product of $\R^2\rtimes SO(2).$ Translate first then rotate. e.g. $g=\left(\begin{pmatrix}
    0\\1
\end{pmatrix},\begin{pmatrix}
    0 & 1\\-1&0
\end{pmatrix}\right)$
and $g'=\left(\begin{pmatrix}
    -1\\0
\end{pmatrix},\frac{1}{\sqrt2}\begin{pmatrix}
    1&1\\-1&1
\end{pmatrix}\right).$ These are the motions of the plane that preserves lengths and angles.

\begin{theorem}
    Let $G$ be a finite group and $G_0=G.$ then construct $G_1=[G_0,G_0],\dots G_i=[G_{i-1},G_{i-1}].$  This series will always terminate, and $G$ is solvable iff $\exists r$ s.t. $G_r=\{e\}.$ \begin{proof}
        Messy, but not hard.
    \end{proof}
\end{theorem}
\begin{definition}[Upper Central Series]
    $Z_0=\{e\},Z_1(G)$ and $Z_1(G)=Z(G)$ then let $Z_2$ be a subgroup of $G$ s.t. $Z_2(G)/Z_1(G)=Z(G/Z_1(G)),$ $Z_i(G)/Z_{i-1}(G)=Z(G/Z_{i-1}(G)).$ This series will always terminate with \begin{align}
        \{e\}=Z_0\triangleleft Z_1\triangleleft Z_2\triangleleft\cdots\triangleleft Z_r=G
    \end{align}
\end{definition}
\begin{definition}[Nilpotent]
    $G$ is \bluebf{nilpotent} if $G$ is solvable and $Z_r(G)=G.$
\end{definition}
\begin{definition}[Lower Central Series]
    $G^{(0)}=G,G^{(1)}=[G,G],G^{(2)}=[G^{(0)},G^{(1)}],G^{(i)}=[G^{(0)},G^{(i-1)}].$   
\end{definition}
\begin{theorem}
    $G$ is solvable iff $\exists r$ s.t. $G^{(r)}=\{e\}.$
\end{theorem}

We have developed the following understanding of groups in order of complexity: \begin{enumerate}
    \item Trivial group $\{e\}$
    \item Cyclic group of prime order $C_p$
    \item Cyclic group $C_n$
    \item Abelian group
    \item $p$-group
    \item Nilpotent group
    \item Solvable group
\end{enumerate}

\begin{definition}[Characteristic]
    A proper subgroup $H<G$ is a \bluebf{characteristic subgroup} if $\varphi(H)=H$ for all $\varphi\in\mathrm{Alt}(G).$ (Note normal subgroups are only required to satisfy this property for inner automorphisms).
\end{definition}
\begin{proposition}
    If $H$ is normal in a characteristic subgroup of $G,$ then $H\trianglelefteq G.$ This is not true without the characteristic property.
\end{proposition}

\subsection{Nilpotent Groups}
Easy: $p$-groups are nilpotent.

(almost) easy: a product of nilpotent groups is nilpotent. (the pieces in the definition of nilpotence work ``component-wise'' in a product).

In particular, product of $p$-groups are nilpotent.

If $P$ is a $p$-group and $Q$ is a $q$-group, in $G=P\times Q,$ $P=P\times\{e\}\in\mathrm{Syl}_p(G)\triangleleft G$ and $Q=\{e\}\times Q\in\mathrm{Syl}_q(G)\triangleleft G.$ Analogously, $G=P_1\times P_2\times\cdots\times P_k$ is nilpotent iff $P_i$ is a $p_i$-group, then the $P_i$s are the Sylow $p_i$-subgroups of $G,$ and each is normal (so it is the only $p_i$ subgroup).

\begin{theorem}
    Suppose $G$ is finite, then the following are equivalent: \begin{enumerate}
        \item $G$ is nilpotent.
        \item If $H<G$ is a proper subgroup, then $H<N_G(H)$ is also a proper subgroup.
        \item If $p\mid |G|$ and $P\in \mathrm{Syl}_p(G),$ then $P$ is normal. Hence, all Sylow $p$-subgroups are normal.
        \item $G\cong P_1\times\cdots\times P_k$ where $P_i\in\mathrm{Syl}_{p_i}(G)$ and $p_1,\cdots,p_k$ are distinct primes. 
    \end{enumerate}
    \begin{proof}[Proof (hint $1\to 2$)]
        If $G$ is abelian, then the proof is trivial. So, we can assume $G$ is not abelian. Otherwise, the proof of the theorem is trivial.
    \end{proof}
\end{theorem}
Consider a finite field $\mathbb F$ and matrices over $\mathbb F$ with the form $\begin{pmatrix}
    1&a&b\\0&1&c\\0&0&1
\end{pmatrix}.$ Suppose two matrices of this form, \begin{align}
    \underbrace{\begin{pmatrix}
    1&a&b\\0&1&c\\0&0&1
\end{pmatrix}}_g\underbrace{\begin{pmatrix}
    1&x&y\\0&1&z\\0&0&1
\end{pmatrix}}_h&=\begin{pmatrix}
    1&a+x&b+az+y\\0&1&c+z\\0&0&1
\end{pmatrix}\\
hg&=\begin{pmatrix}
    1&a+x&b+cx+y\\0&1&c+z\\0&0&1
\end{pmatrix}\\
[g,h]=g\inv h\inv gh&=\begin{pmatrix}
    1&0&x\\0&1&0\\0&0&1
\end{pmatrix}
\end{align}
If $G=\left\{\begin{pmatrix}
    1&*&*\\0&1&*\\0&0&1
\end{pmatrix}\right\},$ then $[G,G]=\left\{\begin{pmatrix}
    1&0&*\\0&1&0\\0&0&1
\end{pmatrix}\right\},$ and $[G,[G,G]]=\{e\}.$ This shows $G$ is solvable. If we extend $n$ to any number beyond $3,$ the same argument holds.

\end{document}

